---
title: "Median Income with IPUMS"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

This script calculates the weighted median income of an arbitrary region (group of counties) using data from IPUMS-USA.

```{r message = FALSE}
library(tidyverse)
library(ipumsr)
library(matrixStats)
```

Read the data files.

```{r}
ddi <- read_ipums_ddi("usa_00006.xml")
data <- read_ipums_micro(ddi)
```

Remove irrelevant data, including wages listed as 99999 and observations under age 25. Then calculate the weighted median income using the weightedMedian() function from the matrixStats library.

```{r}
tx <- data %>% filter(STATEFIP==48, INCWAGE != 0, INCWAGE < 999999, AGE >= 25)
weightedMedian(tx$INCWAGE, tx$PERWT)
```

The FIPS codes listed here represent the counties of the Comptroller's Gulf Coast region. First calculate the total median income of the region, then calculate it by educational attainment: high school or equivalent, associate degree, bachelor's, and master's.

```{r}
gulfcoast <- tx %>% filter(COUNTYFIP %in% c(15,39,71,89,157,167,201,291,321,339,471,473,481))
weightedMedian(gulfcoast$INCWAGE, gulfcoast$PERWT)
gchighsch <- filter(gulfcoast, EDUCD == 62 | EDUCD == 63 | EDUCD == 64)
weightedMedian(gchighsch$INCWAGE, gchighsch$PERWT)
gcass <- filter(gulfcoast, EDUCD == 81 | EDUCD == 82 | EDUCD == 83)
weightedMedian(gcass$INCWAGE, gcass$PERWT)
gcbach <- filter(gulfcoast, EDUCD == 101)
weightedMedian(gcbach$INCWAGE, gcbach$PERWT)
gcmasters <- filter(gulfcoast, EDUCD==114)
weightedMedian(gcmasters$INCWAGE, gcmasters$PERWT)
```
